---
title: "Global emissions trends by gas"
author: "William F. Lamb"
output: word_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../../Results/knitr/") })

---


```{r setup, include=FALSE}

rm(list = ls())
library(tidyverse)
library(ggpubr)
library(patchwork)
library(openxlsx)
library(ggplot2); theme_set(theme_bw())

source("small_figures.R")

#load('../../Data/edgar_essd_data_ghg_gwp_ar6.RData')
#load('../../Data/edgar_essd_data_raw.RData')

load('../../Data/edgar6_v4_data_ghg_gwp_ar6.RData')
load('../../Data/edgar6_v4_data_raw.RData')

load('../../Data/gwps.RData')
load("../../Data/land.RData")

uncertainties <- data.frame(gas=c('CO2 FFI','CO2 LULUCF','CH4','N2O','Fgas'),
                            uncertainty=c(0.08,0.7,0.3,0.6,0.3))

```


```{r waterfall_data_v1, echo=FALSE,warning=FALSE,include=FALSE}

####### calculate total emissions for different gases using different GWPs

waterfall_data <- edgar_raw %>%
  group_by(year,gas,gwp100_ar2,gwp100_ar5,gwp100_ar6) %>%
  summarise(value=sum(value,na.rm=TRUE)) %>%
  mutate(value=value/1e9) %>%
  ungroup()

waterfall_data <- waterfall_data %>%
  mutate(gwp_ar2=gwp100_ar2*value) %>%
  mutate(gwp_ar5=gwp100_ar5*value) %>%
  mutate(gwp_ar6=gwp100_ar6*value) %>%
  group_by(year,gas) %>%
  summarise_at(vars(gwp_ar2:gwp_ar6),sum,na.rm=TRUE)

#### calculate fgas and land emissions

fgas <- waterfall_data %>%
  filter(!gas %in% c("CO2","CH4","N2O")) %>%
  group_by(year) %>%
  summarise_at(vars(gwp_ar2:gwp_ar6),sum,na.rm=TRUE) %>%
  mutate(gas="Fgas") %>%
  select(year,gas,everything())

land_totals <- land %>%
  filter(year>1969) %>%
  filter(year<2020) %>%
  group_by(year) %>%
  summarise(CO2_LULUCF=sum(mean)/1e9)

waterfall_land <- land_totals %>%
  mutate(gas="CO2 LULUCF") %>%
  mutate(gwp_ar2=CO2_LULUCF) %>%
  mutate(gwp_ar5=CO2_LULUCF) %>%
  mutate(gwp_ar6=CO2_LULUCF) %>%
  select(-CO2_LULUCF)

waterfall_data <- waterfall_data %>% filter(gas %in% c("CO2","CH4","N2O"))
waterfall_data <- rbind(waterfall_data,fgas)
waterfall_data <- rbind(waterfall_data,waterfall_land)

waterfall_data <- waterfall_data %>%
  mutate(gas=as.character(gas)) %>%
  mutate(gas=ifelse(gas=="CO2","CO2 FFI",gas))

waterfall_data <- waterfall_data %>%
  filter(year==2019) %>%
  select(-year)

# reorder levels
waterfall_data$gas <- as.factor(waterfall_data$gas)
waterfall_data$gas <-  factor(waterfall_data$gas,levels(waterfall_data$gas)[c(2,3,1,5,4)])
waterfall_data <- waterfall_data %>% arrange(gas)


######### save data #########

save_waterfall_data <- waterfall_data

######### plot different gwp totals in a waterfall


waterfall_data <- gather(waterfall_data,gwp,value,gwp_ar2:gwp_ar6)

waterfall_data <- waterfall_data %>%
  group_by(gwp) %>%
  mutate(end = cumsum(value)) %>%
  mutate(start = c(0, head(end, -1))) %>%
  mutate(id = c(5,4,3,2,1))

### add uncertainties

waterfall_data <- left_join(waterfall_data,uncertainties,by=c("gas"="gas"))
waterfall_data <- waterfall_data %>%
  group_by(gwp) %>%
  mutate(abs_uncertainty=value*uncertainty) %>%
  mutate(uncertainty_start=end-abs_uncertainty) %>%
  mutate(uncertainty_end=end+abs_uncertainty) %>%
  group_by(gwp) %>%
  mutate(total=sum(value))

waterfall_data$gas <- as.factor(waterfall_data$gas)
waterfall_data$gas <-  factor(waterfall_data$gas,levels(waterfall_data$gas)[c(4,5,1,3,2)])

waterfall_data$gwp <- as.factor(waterfall_data$gwp)
waterfall_data$gwp <-  factor(waterfall_data$gwp,levels(waterfall_data$gwp)[c(3,2,1)])


```

```{r waterfall_data_v2, echo=FALSE,warning=FALSE,include=FALSE}

####### calculate total emissions for different gases using different GWPs
# 
# waterfall_data <- edgar_raw %>% 
#   group_by(year,gas,gwp100_ar2,gwp100_ar5,gwp100_ar5_fb,gwp100_ar6) %>% 
#   summarise(value=sum(value,na.rm=TRUE)) %>% 
#   mutate(value=value/1e9) %>%
#   ungroup()
# 
# waterfall_data <- waterfall_data %>% 
#   mutate(gwp_ar2=gwp100_ar2*value) %>% 
#   mutate(gwp_ar5=gwp100_ar5*value) %>% 
#   mutate(gwp_ar5_fb=gwp100_ar5_fb*value) %>% 
#   mutate(gwp_ar6=gwp100_ar6*value) %>% 
#   group_by(year,gas) %>% 
#   summarise_at(vars(gwp_ar2:gwp_ar6),sum,na.rm=TRUE)
# 
# #### calculate fgas and land emissions
# 
# fgas <- waterfall_data %>% 
#   filter(!gas %in% c("CO2","CH4","N2O")) %>% 
#   group_by(year) %>% 
#   summarise_at(vars(gwp_ar2:gwp_ar6),sum,na.rm=TRUE) %>% 
#   mutate(gas="Fgas") %>% 
#   select(year,gas,everything())
# 
# land_totals <- land %>% 
#   filter(year>1969) %>% 
#   filter(year<2020) %>% 
#   group_by(year) %>% 
#   summarise(CO2_LULUCF=sum(mean)/1e9)
# 
# waterfall_land <- land_totals %>% 
#   mutate(gas="CO2 LULUCF") %>% 
#   mutate(gwp_ar2=CO2_LULUCF) %>% 
#   mutate(gwp_ar5=CO2_LULUCF) %>% 
#   mutate(gwp_ar5_fb=CO2_LULUCF) %>% 
#   mutate(gwp_ar6=CO2_LULUCF) %>% 
#   select(-CO2_LULUCF)
# 
# waterfall_data <- waterfall_data %>% filter(gas %in% c("CO2","CH4","N2O"))
# waterfall_data <- rbind(waterfall_data,fgas)
# waterfall_data <- rbind(waterfall_data,waterfall_land)
# 
# waterfall_data <- waterfall_data %>% 
#   mutate(gas=as.character(gas)) %>%
#   mutate(gas=ifelse(gas=="CO2","CO2 FFI",gas))
# 
# waterfall_data <- waterfall_data %>% 
#   filter(year==2019) %>% 
#   select(-year)
# 
# # reorder levels
# waterfall_data$gas <- as.factor(waterfall_data$gas)
# waterfall_data$gas <-  factor(waterfall_data$gas,levels(waterfall_data$gas)[c(2,3,1,5,4)])
# waterfall_data <- waterfall_data %>% arrange(gas)
# 
# 
# ######### save data ######### 
# 
# save_waterfall_data <- waterfall_data
# 
# ######### plot different gwp totals in a waterfall
# 
# 
# waterfall_data <- gather(waterfall_data,gwp,value,gwp_ar2:gwp_ar6)
# 
# waterfall_data <- waterfall_data %>% 
#   group_by(gwp) %>% 
#   mutate(end = cumsum(value)) %>% 
#   mutate(start = c(0, head(end, -1))) %>% 
#   mutate(id = c(5,4,3,2,1))
# 
# ### add uncertainties
# 
# waterfall_data <- left_join(waterfall_data,uncertainties,by=c("gas"="gas"))
# waterfall_data <- waterfall_data %>% 
#   group_by(gwp) %>% 
#   mutate(abs_uncertainty=value*uncertainty) %>% 
#   mutate(uncertainty_start=end-abs_uncertainty) %>% 
#   mutate(uncertainty_end=end+abs_uncertainty) %>% 
#   group_by(gwp) %>% 
#   mutate(total=sum(value))
# 
# waterfall_data$gas <- as.factor(waterfall_data$gas)
# waterfall_data$gas <-  factor(waterfall_data$gas,levels(waterfall_data$gas)[c(4,5,1,3,2)])
# 
# waterfall_data$gwp <- as.factor(waterfall_data$gwp)
# waterfall_data$gwp <-  factor(waterfall_data$gwp,levels(waterfall_data$gwp)[c(4,3,2,1)])


```

```{r waterfall_plot, echo=FALSE,warning=FALSE,include=FALSE,fig.width=3,fig.height=4}

### plot

colours = c("#e78ac3","#fc8d62","#8da0cb","#66c2a5","#a6d854")

p_waterfall <- waterfall_data %>%
  ggplot(., aes(fill = gas)) +
  scale_fill_manual(values=colours) +
  
  geom_rect(aes(
    x = gas,
    xmin = id - 0.48,
    xmax = id + 0.48,
    ymin = start,
    ymax = end
  ),colour = "#737373") +

  geom_rect(aes(
    xmin = id,
    xmax = id,
    ymin = uncertainty_start,
    ymax = uncertainty_end
  ),colour = "#252525") +

  geom_rect(aes(
    xmin = id - 0.2,
    xmax = id + 0.2,
    ymin = uncertainty_start,
    ymax = uncertainty_start
  ),colour = "#252525") +

  geom_rect(aes(
    xmin = id - 0.2,
    xmax = id + 0.2,
    ymin = uncertainty_end,
    ymax = uncertainty_end
  ),colour = "#252525") +

   facet_grid(~gwp,switch="both") +

   geom_text(data=waterfall_data %>% filter(gas=="CH4"),aes(x=gas,y=62,label=paste(round(total,0),  "Gt")),size=3.5,colour="#252525") +

   theme_bw() +
   scale_y_continuous(breaks=c(0,10,20,30,40,50,60)) +
   scale_x_discrete(labels=c("","","","","")) +
   theme(
     legend.position = "none",
     title = element_blank(),
     axis.title.y = element_blank(),
     axis.ticks.y = element_blank(),
     axis.text.y = element_blank(),
     axis.ticks.x = element_blank(),
     panel.grid.minor.x = element_blank(),
     panel.grid.minor.y = element_blank(),
     panel.grid.major.x = element_blank(),
     panel.spacing = unit(0, "mm"),
     strip.background = element_blank())

p_waterfall

```


```{r gas_trend, echo=FALSE,warning=FALSE,include=FALSE,fig.width=8,fig.height=4}


gas_trend <- edgar_ghg %>% 
  filter(year>1989) %>% 
  filter(year<2020) %>% 
  group_by(year) %>% 
  summarise(CO2=sum(CO2,na.rm=T)/1e9,CH4=sum(CH4,na.rm=T)/1e9,N2O=sum(N2O,na.rm=T)/1e9,Fgas=sum(Fgas,na.rm=T)/1e9)

### join land use data

gas_trend <- left_join(gas_trend,land_totals,by=c("year"="year"))
gas_trend<- gather(gas_trend,gas,value,-year) 
gas_trend <- gas_trend %>% 
  mutate(gas=ifelse(gas=="CO2","CO2 FFI",gas)) %>% 
  mutate(gas=ifelse(gas=="CO2_LULUCF","CO2 LULUCF",gas))

gas_trend$gas <- as.factor(gas_trend$gas)
gas_trend$gas <-  factor(gas_trend$gas,levels(gas_trend$gas)[c(4,5,1,3,2)])

### get the shares and totals in each time period

shares <- gas_trend %>% 
  filter(year %in% c(1990,2000,2010,2019)) %>% 
  group_by(year) %>% 
  mutate(totals=sum(value)) %>% 
  ungroup() %>% 
  group_by(year,gas) %>% 
  mutate(fractions=(value/totals)*100)

shares <- locate_shares(shares,"gas",4)

line_data <- data.frame(x=c(1990,1990,2000,2000,2010,2010,2019,2019),y=c(0,60))

### get the growth rates between time periods
rates <- gas_trend %>% 
  filter(year %in% c(1990,1999,2000,2009,2010,2019)) %>% 
  group_by(year) %>% 
  summarise(value=sum(value)) %>% 
  mutate(total_rate=NA)

rates$total_rate[rates$year==1990] = ((rates$value[rates$year==1999]/rates$value[rates$year==1990])^(1/(1999-1990))-1)*100
rates$total_rate[rates$year==2000] = ((rates$value[rates$year==2009]/rates$value[rates$year==2000])^(1/(2009-2000))-1)*100
rates$total_rate[rates$year==2010] = ((rates$value[rates$year==2019]/rates$value[rates$year==2010])^(1/(2019-2010))-1)*100
  

p_gas_totals <- gas_trend %>% 
  ggplot(.,aes(x=year,y=value,fill=gas)) +
  geom_area(colour="#737373") +
  
  geom_line(inherit.aes = FALSE,data=line_data,aes(x=x,y=y,group=x),alpha=0.3,linetype="dashed") +
  
  #### text with the shares
  geom_text(data=shares,aes(x=year+0.25,y=location,label=paste(round(fractions,0),"%",sep="")),size=3.5,colour="#252525",hjust=0)+
  
  #### text with the rates
  geom_text(data=rates %>% filter(!is.na(total_rate)),inherit.aes = FALSE,
                            aes(x=ifelse(year==2010,year+4.5,year+5),y=60,
                                label=paste0("+",round(total_rate,1),"%")),size=3.5,colour="#252525") +
  
  #### text with the totals
  geom_text(data=shares %>% filter(gas=="Fgas"),aes(x=year,y=62,label=paste(round(totals,0),  "Gt")),size=3.5,colour="#252525")+
  
  ylab('Greenhouse gas emissions (GtCO2eq/yr)') +
  labs(fill="Gas") +
  scale_y_continuous(breaks=c(0,10,20,30,40,50,60)) +
  scale_x_continuous(breaks=c(1990,2000,2010,2019)) +
  scale_fill_manual(values=c("#e78ac3","#fc8d62","#8da0cb","#66c2a5","#a6d854")) +
  big_trend_theme +
  theme(
    legend.position="none",    
    legend.background = element_rect(linetype = 1, size = 0.5, colour = "#525252"),
    legend.margin = margin(l=5,t=5,b=5,r=10,unit="pt"),
    legend.text = element_text(size=8),
    legend.title = element_text(size=8)) +
  guides(fill=guide_legend(nrow=1,byrow=TRUE),
    shape = guide_legend(override.aes = list(size = 0.5)),
    color = guide_legend(override.aes = list(size = 0.5)))+
  ggtitle("a. Total anthropogenic GHG emissions 1990-2019")

p_gas_totals

```

```{r ghgs_no_gwp, echo=FALSE,warning=FALSE,include=FALSE,fig.width=8,fig.height=2}


gas_trend_nogwp <- edgar_raw %>%
  filter(year<2020) %>% 
  filter(gas %in% c("CO2","CH4","N2O")) %>% 
  group_by(year,gas) %>%
  summarise(value=sum(value,na.rm=TRUE))

fgas_trend <- edgar_ghg %>% 
  filter(year<2020) %>% 
  group_by(year) %>% 
  summarise(value=sum(Fgas,na.rm=TRUE)) %>% 
  mutate(gas="Fgas")

gas_trend_nogwp <- rbind(gas_trend_nogwp,land_totals %>% select(year,value=CO2_LULUCF) %>% mutate(gas="CO2 LULUCF") %>% mutate(value=value*1e9))
gas_trend_nogwp <- rbind(gas_trend_nogwp,fgas_trend)
gas_trend_nogwp <- gas_trend_nogwp %>% mutate(gas=ifelse(gas=="CO2","CO2 FFI",gas))


# Normalise to 1 in 1990
gas_trend_nogwp <- gas_trend_nogwp %>%
  filter(year>1989) %>%
  group_by(gas) %>%
  mutate(value=value/(value[year==1990]))

# calculate error range
gas_trend_nogwp <- left_join(gas_trend_nogwp,uncertainties,by = "gas")

gas_trend_nogwp <- gas_trend_nogwp %>%
  mutate(high = value+((uncertainty*value))) %>%
  mutate(low = value-((uncertainty*value))) %>%
  select(-uncertainty)

#manually map colours
gas_trend_nogwp$gas <- as.factor(gas_trend_nogwp$gas)
gas_trend_nogwp$gas <- factor(gas_trend_nogwp$gas,levels=levels(gas_trend_nogwp$gas)[c(2,3,1,5,4)])
colours = c("#a6d854","#66c2a5","#8da0cb","#fc8d62","#e78ac3")


p_gas_trends <- gas_trend_nogwp %>%
  ggplot(.,aes(x=year,y=value)) +
  geom_hline(yintercept = 1,color="#636363") +
  geom_line(aes(color=gas)) +
  geom_ribbon(data=gas_trend_nogwp,aes(ymin=low,ymax=high,fill=gas),alpha=0.3)+
  scale_x_continuous(breaks=c(1990,2000,2010)) +
  scale_fill_manual(values=colours) +
  scale_color_manual(values=colours) +
  theme_bw() +
  ylab(str_wrap('Emissions (original units, relative to 1990)',width=26)) +
  facet_wrap(.~gas,nrow=1) +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.title.y = element_text(margin=margin(0,10,0,0)),
        text = element_text(size=11),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())



p_gas_trends


```

```{r ghgs_no_gwp_scales, echo=FALSE,warning=FALSE,include=FALSE,fig.width=8,fig.height=2}


p1 <- gas_trend_nogwp %>%
  filter(gas!="Fgas") %>%
  ggplot(.,aes(x=year,y=value)) +
  geom_hline(yintercept = 1,color="#636363") +
  geom_line(aes(color=gas),width=2) +
  geom_ribbon(data=gas_trend_nogwp %>% filter(gas!="Fgas"),aes(ymin=low,ymax=high,fill=gas),alpha=0.3)+
  scale_x_continuous(breaks=c(1990,2000,2010)) +
  scale_fill_manual(values=colours) +
  scale_color_manual(values=colours) +
  theme_bw() +
  ylab('GHG Emissions\n(relative to 1990)') +
  facet_wrap(.~gas,nrow=1) +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.title.y = element_text(margin=margin(0,10,0,0)),
        text = element_text(size=11),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  ggtitle("b. Trends in global greenhouse gas emissions and their uncertainties, relative to 1990")

p2 <- gas_trend_nogwp %>%
  filter(gas=="Fgas") %>%
  ggplot(.,aes(x=year,y=value)) +
  geom_hline(yintercept = 1,color="#636363") +
  geom_line(aes(color=gas),width=2) +
  geom_ribbon(data=gas_trend_nogwp %>% filter(gas=="Fgas"),aes(ymin=low,ymax=high,fill=gas),alpha=0.3)+
  scale_x_continuous(breaks=c(1990,2000,2010)) +
  #scale_y_continuous(limits = c(-3,7),breaks=c(-2,1,2,4)) +
  scale_fill_manual(values="#e78ac3") +
  scale_color_manual(values="#e78ac3") +
  theme_bw() +
  facet_wrap(.~gas,nrow=1) +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        text = element_text(size=11),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),plot.background = element_blank())


p_gas_trends_scaled <- p1 + p2 + plot_layout(width = c(4,1))
p_gas_trends_scaled

```

```{r gas_trends_spm, echo=FALSE,warning=FALSE,fig.width=8,fig.height=6,fig.path="../../Results/Plots/",dev=c('png','pdf')}


#########

(p_gas_totals + p_waterfall + plot_layout(width = c(3, 1))) / p_gas_trends_scaled + plot_layout(height = c(3,1))
#gases <- (p_gas_totals + p_waterfall + plot_layout(width = c(2.5, 1))) / p_gas_trends_scaled + plot_layout(height = c(3,1))
#gases

##### write data

wb <- createWorkbook(title = paste("ipcc_ar6_figure_spm1_gases"))

addWorksheet(wb,"info")
addWorksheet(wb,"gas_trend")
addWorksheet(wb,"gwp_waterfall")
addWorksheet(wb,"gas_trend_normalised")

info = data.frame(x=c("Author","Last update","Code"),y=c("William F. Lamb",as.character(Sys.time()),"https://github.com/mcc-apsis/AR6-Emissions-trends-and-drivers/blob/master/R/Analysis%20and%20figures/emissions_by_gas.Rmd"))

writeData(wb, sheet="info",info,colNames=F,rowNames=F)
writeData(wb, sheet = "gas_trend", spread(gas_trend,year,value), colNames = T, rowNames = F)
writeData(wb, sheet = "gwp_waterfall", waterfall_data, colNames = T, rowNames = F)
writeData(wb, sheet = "gas_trend_normalised", gas_trend_nogwp, colNames = T, rowNames = F)

saveWorkbook(wb,paste0("../../Results/Plot data/ipcc_ar6_figure_spm1_gases.xlsx"),overwrite=T)
