---
title: "Emissions trends for WGII Regions"
author: "William F. Lamb (Working Group III Ch2; Lamb@mcc-berlin.net)"
output: 
  word_document:
    toc: yes
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../../Results/knitr/") })
  
---

# Introduction

## Description
These figures depict emissions trends for the WGII regions using data compiled by WGIII. Code for producing them can be found here: https://github.com/mcc-apsis/AR6-Emissions-trends-and-drivers/blob/master/R/wg2_regions.Rmd. The figures are available as PNG and PDF files here: https://github.com/mcc-apsis/AR6-Emissions-trends-and-drivers/tree/master/Results/Plots/WG2_regions
An excel file with the data compiled for each figure is available on request. 

## Figure caption (replace #region as appropriate)
Title: Greenhouse gas emissions trends for (#region). 
Caption: Panel a depicts per capita greenhouse gas emissions by region and growth from 1990-2018. Panel b depicts total greenhouse gas emissions by region since 1990. Panel c depicts the absolute emissions growth between 2010 and 2018 for the highest emitting countries within (#region). Panel d depicts total (#region) emissions since 1990, broken down by greenhouse gas (left) and sector (right). 100 year global warming potentials consistent with WGI estimates are used. Emissions data are from Crippa et al. (2019), compiled by Ch2 of WGIII. Estimates of land-use change CO2 are not included.

## Modifications and requests
I will do some final manual modifications before the final draft (e.g. tidying the figure legend). Alternative presentations of the data may also be possible, on request.

## Data citation (TO BE UPDATED BEFORE FINAL DRAFT)
Crippa, M., Oreggioni, G., Guizzardi, D., Muntean, M., Schaaf, E., Lo Vullo, E., … Vignati, E. (2019). Fossil CO2 and GHG emissions of all world countries - 2019 Report. Luxembourg. https://doi.org/10.2760/687800

.ris file: https://github.com/mcc-apsis/AR6-Emissions-trends-and-drivers/blob/master/Results/Plots/WG2_regions/Crippa%20et%20al.%202019.ris

# Script setup

```{r setup, include=FALSE}

rm(list = ls())
library(tidyverse)
library(ggpubr)
library(zoo)
library(ggmap)
library(maps)
library(ggrepel)
library(patchwork)

load('../../Data/edgar6_v2_data_ghg_gwp_ar6.RData')
load('../../Data/gwps.RData')
load('../../Data/WDI_gdp_pop.RData')
load('../../Data/land.RData')


# set palette
ggplot <- function(...) ggplot2::ggplot(...) + scale_color_brewer(palette="Set2") + scale_fill_brewer(palette="Set2")


uncertainties <- data.frame(gas=c('CO2 FFI','CO2 FOLU','CH4','N2O','Fgas','GHG'),
                            uncertainty=c(0.08,0.5,0.2,0.6,0.2,0.1))

wb <- openxlsx::createWorkbook(title = paste0("ipcc_wg2_region_data_",Sys.Date()))

edgar_ghg <- edgar_ghg %>% 
  mutate(chapter_title=ifelse(chapter_title=="AFOLU","Agriculture",chapter_title))

```

# Data
## Join WG2 regions to the WG3 data

```{r, wg2_regions, include=FALSE, echo=FALSE}

regions <- openxlsx::read.xlsx("../../Data/Codes and classifications/Regions_Countries_WGII.xlsx")

regions <- regions %>% 
  select(wg2_region=WGII_Region_checked,ISO=ISO_A3)


data <- edgar_ghg
data <- left_join(data,regions,by = c("ISO"))

not_joined <- anti_join(data,regions,by = c("ISO"))
not_joined <- not_joined %>% 
  ungroup() %>% 
  filter(year==2018) %>% 
  select(country,ISO) %>% 
  distinct()

data <- data %>% 
  mutate(wg2_region=ifelse(wg2_region=="Central and South America","Central and\nS. America",wg2_region))


```

## Per capita GHG

``` {r per_capita_boxplots, echo=FALSE,warning=FALSE,fig.width=8,fig.height=4}


boxplot_data <- data %>% 
  filter(year %in% c(1990,2019)) %>% 
  group_by(wg2_region,year,ISO,country) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE)) %>% 
  mutate(GHG=GHG) %>% 
  filter(!is.na(wg2_region))


#boxplot_data <- left_join(boxplot_data,basic %>% select(ISO,year=Year,pop_UN),by = c("ISO", "year"))
boxplot_data <- left_join(boxplot_data,wdi_data_gdp_pop %>% select(ISO=iso3c,year,pop=population),by = c("ISO", "year"))


### Taiwan and some other missing from boxplots

not_joined <- anti_join(boxplot_data,wdi_data_gdp_pop %>% select(ISO=iso3c,year,pop=population),by = c("ISO", "year"))

boxplot_data <- boxplot_data %>% 
  mutate(GHGpc=(GHG)/pop) %>% 
  ungroup()

boxplot_data$year = as.factor(boxplot_data$year)
boxplot_data$year = factor(boxplot_data$year,levels=levels(boxplot_data$year)[c(2,1)])

boxplot_data <- boxplot_data %>% 
  mutate(wg2_region = as.factor(wg2_region))

## calculate region averages
boxplot_data_regions <- boxplot_data %>% 
  group_by(wg2_region,year) %>% 
  summarise(GHG_region=sum(GHG,na.rm = TRUE),pop_region=sum(pop,na.rm = TRUE))

boxplot_data_regions <- boxplot_data_regions %>% 
  mutate(GHG_pc_region=GHG_region/pop_region)

boxplot_data <- left_join(boxplot_data,boxplot_data_regions,by = c("wg2_region", "year"))

boxplot_data_regions <- boxplot_data_regions %>% 
  mutate(GHGpc = GHG_pc_region)

p1 <- boxplot_data %>% 
  ggplot(.,aes(x=wg2_region,y=GHGpc,fill=year,group=year)) +
  geom_point(shape=21,color="#636363",position=position_dodge(width=0.5),alpha=0.5) +
  geom_point(data=boxplot_data_regions,color="#636363",shape=23,size=5,position=position_dodge(width=0.5)) +
  #geom_boxplot(position="dodge") +
  coord_flip() +
  theme_bw() +
  guides(fill = guide_legend(reverse=T)) +
  scale_y_continuous(limits=c(0,50)) +
  scale_x_discrete(limits=rev(levels(boxplot_data$wg2_region))) +
  ylab("GHG Emissions per capita (tCO2eq/capita)") +
  theme(legend.position=c(0.9,0.15),
        legend.title = element_blank(),
        axis.title.y = element_blank(),
        strip.text = element_text(margin = margin(.32, 0, .32, 0, "cm"))) +
  ggtitle("a. Regional per capita emissions")



p1
```

## GHG trend


```{r, plot_trends, fig.width=4,fig.height=4,include=FALSE, echo=FALSE}


trend_data <- data %>% 
  filter(year>1989) %>% 
  group_by(wg2_region,year) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE)) %>% 
  mutate(GHG=GHG/1e9) %>% 
  filter(!is.na(wg2_region))

# trend_data <- trend_data %>%
#   mutate(wg2_region=as.factor(wg2_region))
# trend_data$wg2_region <- fct_relevel(trend_data$wg2_region,"Africa")

trend_colours <- c("#66c2a5","#fc8d62","#8da0cb","#e78ac3","#a6d854","#e5c494","#b3b3b3")

p2 <- trend_data %>% 
  ggplot(.,aes(x=year,y=GHG,color=wg2_region)) +
  theme_bw() +
  geom_path(data=trend_data %>% filter(wg2_region!="Africa"),size=1.5) +
  geom_path(data=trend_data %>% filter(wg2_region=="Africa"),size=1.5) +
  geom_point(data=trend_data %>% filter(wg2_region!="Africa"),color="white",size=0.5) +
  geom_point(data=trend_data %>% filter(wg2_region=="Africa"),color="white",size=0.5) +
  geom_point(data=trend_data %>% filter(wg2_region!="Africa") %>% filter(year %in% c(1990,1995,2000,2005,2010,2015,2019)),aes(color=wg2_region),shape=21,size=3,stroke=1,fill="white") +
  scale_color_manual(values=trend_colours) +
  geom_point(data=trend_data %>% filter(wg2_region=="Africa") %>% filter(year %in% c(1990,1995,2000,2005,2010,2015,2019)),aes(color=wg2_region),shape=21,size=3,stroke=1,fill="white") +
  scale_color_manual(values=trend_colours) +
  geom_text_repel(data=trend_data %>% filter(year==2019),aes(x=year+1,y=GHG,label=wg2_region,color=wg2_region),
    nudge_x      = 1,
    direction    = "y",
    hjust        = 0,
    segment.colour = NA
  ) +
  
  scale_x_continuous(breaks=c(1990,2000,2010,2019),limits = c(1990,2030)) +
  ggtitle("b. Regional emissions trends") +
  ylab("GHG Emissions (Gt CO2eq/year)") +
  theme(legend.position="none",
        axis.title.x = element_blank())

p2

```

## Country trends

```{r country_plot, include=FALSE, echo=FALSE}

country_plot <- function(data,region) {
  

growth_data <- data %>% 
  filter(wg2_region==region) %>% 
  filter(year %in% c(1990,2019)) %>% 
  group_by(country,year,wg2_region) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE)) %>% 
  mutate(GHG=GHG/1e6)

growth_data <- spread(growth_data,year,GHG)

growth_data <- growth_data %>% 
  arrange(desc(`2019`)) %>% 
  head(15)

growth_data$country <- as.factor(growth_data$country)
growth_data$country <- fct_reorder(growth_data$country,growth_data$`2019`)

growth_data <- gather(growth_data,year,value,`1990`:`2019`)


growth_data$year = as.factor(growth_data$year)
growth_data$year = factor(growth_data$year,levels=levels(growth_data$year)[c(2,1)])


p <- growth_data %>% 
  ggplot(.,aes(x=country,y=value,fill=year)) +
  geom_bar(stat='identity',width=0.7,position=position_dodge(width=0.7),color="#636363") +
  theme_bw() +
  guides(fill = guide_legend(reverse=T)) +
  coord_flip() +
  ylab("GHG Emissions (MtCO2eq)") +
  theme(axis.title.y = element_blank(),
        legend.position=c(0.9,0.15),
        legend.title = element_blank())


#growth_data <- spread(growth_data,year,GHG)
# growth_data <- growth_data %>% 
#   mutate(growth=`2019`-`2010`) %>% 
#   arrange(desc(`2019`))
# 
# last_country <- growth_data %>% head(15) %>% ungroup() %>% arrange(desc(growth)) %>% slice(n())
# axis_limit <- growth_data %>% head(15) %>% ungroup() %>% arrange(growth) %>% slice(n())
# 
# text_thing <- data.frame(x=last_country$country,y=axis_limit$growth)
#   
#
# p <- growth_data %>% 
#   head(15) %>% 
#   ggplot(.,aes(x=reorder(country,growth),y=growth)) +
#   geom_bar(stat='identity',fill="#969696") +
#   theme_bw() +
#   geom_text(data=text_thing,aes(x=x,y=y,label="Years: 2010-2019"),hjust=1) +
#   coord_flip() +
#   ylab("GHG Emissions growth (MtCO2)") +
#   theme(axis.title.y = element_blank())
# p

  return(list("plot"=p,"data"=growth_data))
}


```

## GHGs by gas and sector

```{r gas_sector_plot, echo=FALSE,warning=FALSE,fig.width=6,fig.height=4}

gas_sector_plot <- function(data,region) {
  
  ## gases
  
  bar_data_gas <- data %>% 
    filter(wg2_region==region) %>% 
    filter(year %in% c(1990,2000,2010,2019)) %>% 
    group_by(year) %>% 
    summarise_at(vars(CO2,CH4,N2O,Fgas),sum,na.rm=TRUE)
  
  bar_data_gas <- gather(bar_data_gas,gas,value,CO2:Fgas)
  
  bar_data_gas <- bar_data_gas %>% 
    mutate(value=value/1e9) %>% 
    mutate(year=as.factor(year))
  
  p1 <- bar_data_gas %>% ggplot(.,aes(x=year,y=value,fill=gas)) +
    geom_bar(stat="identity",colour="#636363") +
    theme_bw() +
    guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
    #guides(fill = guide_legend(reverse=TRUE)) + 
    #scale_x_discrete(limits = rev(levels(bar_data_gas$year))) +
    theme(axis.title.x = element_blank(),
          legend.position="bottom",
          legend.title=element_blank()) +
    ylab("GHG Emissions (Gt CO2eq/year)")
  
  ## sectors
  
  bar_data_sector <- data %>% 
    filter(wg2_region==region) %>% 
    filter(year %in% c(1990,2000,2010,2019)) %>% 
    group_by(year,chapter_title) %>%
    summarise(value=sum(GHG,na.rm=TRUE))
  
  bar_data_sector <- bar_data_sector %>% 
    ungroup() %>% 
    mutate(value=value/1e9) %>% 
    mutate(year=as.factor(year))
  
  p2 <- bar_data_sector %>% ggplot(.,aes(x=year,y=value,fill=chapter_title)) +
    geom_bar(stat="identity",colour="#636363") +
    theme_bw() +
    guides(fill=guide_legend(nrow=3,byrow=TRUE)) +
    #guides(fill = guide_legend(reverse=TRUE)) + 
    #scale_x_discrete(limits = rev(levels(bar_data_sector$year))) +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position="bottom",
          legend.title = element_blank(),plot.background = element_blank()) +
    ylab("GHG Emissions (Gt CO2eq/year)")
  
  combined_data <- rbind(bar_data_gas %>% mutate(plot="gas") %>% select(plot,year,var=gas,value),
                         bar_data_sector %>% mutate(plot="sector") %>% select(plot,year,var=chapter_title,value))
  
  combined_data <- spread(combined_data,year,value)
  
  return(list("gas_plot"=p1,"sector_plot"=p2,"data"=combined_data))
}

  
```

``` {r save}

openxlsx::addWorksheet(wb,"Region per capita")
openxlsx::writeData(wb, sheet = "Region per capita",boxplot_data, colNames = T, rowNames = F)
openxlsx::addWorksheet(wb,"Region trend")
openxlsx::writeData(wb, sheet = "Region trend",spread(trend_data,year,GHG), colNames = T, rowNames = F)

```


# Africa

```{r Africa,echo=FALSE,warning=FALSE,fig.width=10,fig.height=8,fig.path="../../Results/Plots/WG2_regions/",dev=c('png','pdf')}

region="Africa"

countries <- country_plot(data,region)
gas_sector <- gas_sector_plot(data,region)

countries$plot <- countries$plot + ggtitle("c. Country GHG emissions (Africa)")
gas_sector$gas_plot <- gas_sector$gas_plot + ggtitle("d. Total GHG emissions by gas and sector (Africa)")

p4 <- (gas_sector$gas_plot + gas_sector$sector_plot) 

wrap_plots(p1,p2,countries$plot,p4)

openxlsx::addWorksheet(wb,paste0(region," emissions growth"))
openxlsx::addWorksheet(wb,paste0(region," gas and sector"))
openxlsx::writeData(wb, sheet = paste0(region," emissions growth"),countries$data, colNames = T, rowNames = F)
openxlsx::writeData(wb, sheet = paste0(region," gas and sector"),gas_sector$data, colNames = T, rowNames = F)

```

# Asia

```{r Asia, echo=FALSE,warning=FALSE,fig.width=10,fig.height=8,fig.path="../../Results/Plots/WG2_regions/",dev=c('png','pdf')}

region="Asia"


countries <- country_plot(data,region)
gas_sector <- gas_sector_plot(data,region)

countries$plot <- countries$plot + ggtitle("c. Country emissions growth (Asia)")
gas_sector$gas_plot <- gas_sector$gas_plot + ggtitle("d. Emissions by gas and sector (Asia)")

p4 <- (gas_sector$gas_plot + gas_sector$sector_plot) 

wrap_plots(p1,p2,countries$plot,p4)

openxlsx::addWorksheet(wb,paste0(region," emissions growth"))
openxlsx::addWorksheet(wb,paste0(region," gas and sector"))
openxlsx::writeData(wb, sheet = paste0(region," emissions growth"),countries$data, colNames = T, rowNames = F)
openxlsx::writeData(wb, sheet = paste0(region," gas and sector"),gas_sector$data, colNames = T, rowNames = F)

```

# Australasia

```{r Australasia,echo=FALSE,warning=FALSE,fig.width=10,fig.height=8,fig.path="../../Results/Plots/WG2_regions/",dev=c('png','pdf')}

region="Australasia"


countries <- country_plot(data,region)
gas_sector <- gas_sector_plot(data,region)

countries$plot <- countries$plot + ggtitle("c. Country emissions growth (Australasia)")
gas_sector$gas_plot <- gas_sector$gas_plot + ggtitle("d. Emissions by gas and sector (Australasia)")

p4 <- (gas_sector$gas_plot + gas_sector$sector_plot) 

wrap_plots(p1,p2,countries$plot,p4)

openxlsx::addWorksheet(wb,paste0(region," emissions growth"))
openxlsx::addWorksheet(wb,paste0(region," gas and sector"))
openxlsx::writeData(wb, sheet = paste0(region," emissions growth"),countries$data, colNames = T, rowNames = F)
openxlsx::writeData(wb, sheet = paste0(region," gas and sector"),gas_sector$data, colNames = T, rowNames = F)

```

# Central and South America

```{r Central and South America,echo=FALSE,warning=FALSE,fig.width=10,fig.height=8,fig.path="../../Results/Plots/WG2_regions/",dev=c('png','pdf')}

region="Central and\nS. America"


countries <- country_plot(data,region)
gas_sector <- gas_sector_plot(data,region)

countries$plot <- countries$plot + ggtitle("c. Country emissions growth (C.& S. America)")
gas_sector$gas_plot <- gas_sector$gas_plot + ggtitle("d. Emissions by gas and sector (C.& S. America)")

p4 <- (gas_sector$gas_plot + gas_sector$sector_plot)

wrap_plots(p1,p2,countries$plot,p4)

openxlsx::addWorksheet(wb,"C. & S. Am. emissions growth")
openxlsx::addWorksheet(wb,"C. & S. Am. gas and sector")
openxlsx::writeData(wb, "C. & S. Am. emissions growth",countries$data, colNames = T, rowNames = F)
openxlsx::writeData(wb, "C. & S. Am. gas and sector",gas_sector$data, colNames = T, rowNames = F)

```

# Europe

```{r Europe, echo=FALSE,warning=FALSE,fig.width=10,fig.height=8,fig.path="../../Results/Plots/WG2_regions/",dev=c('png','pdf')}

region="Europe"

countries <- country_plot(data,region)
gas_sector <- gas_sector_plot(data,region)

countries$plot <- countries$plot + ggtitle("c. Country emissions growth (Europe)")
gas_sector$gas_plot <- gas_sector$gas_plot + ggtitle("d. Emissions by gas and sector (Europe)")

p4 <- (gas_sector$gas_plot + gas_sector$sector_plot)

wrap_plots(p1,p2,countries$plot,p4)

openxlsx::addWorksheet(wb,paste0(region," emissions growth"))
openxlsx::addWorksheet(wb,paste0(region," gas and sector"))
openxlsx::writeData(wb, sheet = paste0(region," emissions growth"),countries$data, colNames = T, rowNames = F)
openxlsx::writeData(wb, sheet = paste0(region," gas and sector"),gas_sector$data, colNames = T, rowNames = F)

```

# North America

```{r North America, echo=FALSE,warning=FALSE,fig.width=10,fig.height=8,fig.path="../../Results/Plots/WG2_regions/",dev=c('png','pdf')}

region="North America"

countries <- country_plot(data,region)
gas_sector <- gas_sector_plot(data,region)

countries$plot <- countries$plot + ggtitle("c. Country emissions growth (North America)")
gas_sector$gas_plot <- gas_sector$gas_plot + ggtitle("d. Emissions by gas and sector (North America)")

p4 <- (gas_sector$gas_plot + gas_sector$sector_plot)

wrap_plots(p1,p2,countries$plot,p4)

openxlsx::addWorksheet(wb,paste0(region," emissions growth"))
openxlsx::addWorksheet(wb,paste0(region," gas and sector"))
openxlsx::writeData(wb, sheet = paste0(region," emissions growth"),countries$data, colNames = T, rowNames = F)
openxlsx::writeData(wb, sheet = paste0(region," gas and sector"),gas_sector$data, colNames = T, rowNames = F)


```

# Small Islands

```{r Small Islands,echo=FALSE,warning=FALSE,fig.width=10,fig.height=8,fig.path="../../Results/Plots/WG2_regions/",dev=c('png','pdf')}

region="Small Islands"

countries <- country_plot(data,region)
gas_sector <- gas_sector_plot(data,region)

countries$plot <- countries$plot + ggtitle("c. Country emissions growth (Small Islands)")
gas_sector$gas_plot <- gas_sector$gas_plot + ggtitle("d. Emissions by gas and sector (Small Islands)")

p4 <- (gas_sector$gas_plot + gas_sector$sector_plot)

wrap_plots(p1,p2,countries$plot,p4)

openxlsx::addWorksheet(wb,paste0(region," emissions growth"))
openxlsx::addWorksheet(wb,paste0(region," gas and sector"))
openxlsx::writeData(wb, sheet = paste0(region," emissions growth"),countries$data, colNames = T, rowNames = F)
openxlsx::writeData(wb, sheet = paste0(region," gas and sector"),gas_sector$data, colNames = T, rowNames = F)


```

```{r write,echo=FALSE,warning=FALSE}

openxlsx::saveWorkbook(wb,paste0("../../Results/Data/ipcc_wg2_region_data_",Sys.Date(),".xlsx"),overwrite=T)

```


### Contributions by region



```{r contributions,echo=FALSE,warning=FALSE}

contributions <- data %>% 
  filter(year==2019) %>% 
  group_by(wg2_region) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE))

total <- edgar_ghg %>% 
  group_by(year) %>% 
  summarise(GHG=sum(GHG,na.rm=TRUE)) %>% 
  filter(year==2019)

contributions <- contributions %>% 
  mutate(GHG_total = total$GHG[1]) %>% 
  mutate((fraction=GHG/GHG_total)*100)

```














