---
title: "Emissions trends and drivers: sectoral trends"
author: "William F. Lamb, Jan C. Minx"
output: 
  word_document:
    toc: yes
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../Results") })
---
# Summary
This document provides an overview and breakdown of emissions trends by IPCC chapter. We have compiled a consistent allocation of emissions categories to sectors using the EDGAR emissions database. We propose to use this allocation and the following figures across the main sectoral chapters of AR6.

The following supporting information is available online: 

* our code in R
* the sector categorisations
* the figure files
* the data files for each figure
* this document and any updates

Please find these at: https://github.com/mcc-apsis/AR6-Emissions-trends-and-drivers

Latest update: `r format(Sys.time(), '%d %B, %Y, %H:%M')`

**Author information**

William F. Lamb - Mercator Research Institute on Global Commons and Climate Change (MCC); School of Earth and Environment, University of Leeds; Contributing Author in Chapter 2 of the Working Group III Report (Emissions Trends and Drivers); lamb@mcc-berlin.net

Jan C. Minx - Mercator Research Institute on Global Commons and Climate Change (MCC); School of Earth and Environment, University of Leeds; Lead Author in Chapter 2 of the Working Group III Report (Emissions Trends and Drivers); minx@mcc-berlin.net


```{r setup, include=FALSE}

rm(list = ls())
library(tidyverse)
library(ggpubr)
library(gganimate)
library(xlsx)
library(flextable)

```


```{r edgar_prep,include=FALSE}

sector=8
gas="GHG"
time_start=2010
category="chapter"
rate_or_abs="rate"


library(openxlsx)

######### load edgar data ######### 

load('../Data/edgar.RData')

edgar_emissions_plot <- function(time_start,category,sector,gas,rate_or_abs,title) {
  
  ##############  calculate growth rates in given time period ############## 
  
  #sector_code <- edgar_GHG %>%
  #  filter_at(vars(category),all_vars(.==sector))
  #sector_code <- substr(sector_code$sector_code[[1]],1,as.numeric(str_sub(category,-1)))
  
  rates <- edgar_GHG %>% 
    filter_at(vars(category),all_vars(.==sector)) %>% 
    filter(!is.na(WB.income)) %>% 
    filter(Year>=time_start & Year<2018) %>% 
    group_by(Country,Year,WB.income) %>% 
    summarise_at(gas,sum,na.rm=TRUE) %>% 
    ungroup() %>% 
    rename_at(gas,funs(sprintf('value')))
  
  ####### CHECK DATA HERE FOR ZEROS AND NANS ####### 
  
  
  # what is 80% of the emissions?
  blarg <- rates %>%
    filter(Year==2017)
  threshold <- sum(blarg$value,na.rm=T)*0.8
  
  ## remove countries below absolute emissions threshold
  
  rates <- rates %>% 
    group_by(Country) %>% 
    mutate(rate=(last(value)/first(value))^(1/(last(Year)-time_start+1))-1) %>% 
    mutate(abs_growth=last(value)-first(value)) %>%
    filter(Year==2017) %>% 
    na.omit() %>% 
    arrange(desc(value))
  
  
  rates$cutoff <- 0
  z <- 0
  for (i in 1:length(rates$Country)){
    z <- z + rates$value[i]
    if (z > threshold) {
      break
    }
    rates$cutoff[i] <- 1
  }
  
  country_list <- rates %>% select(Country,cutoff)
  
  #### calculate growth rates for countries below threshold, grouped by region ####
  
  blarg <- edgar_GHG %>% 
    filter_at(vars(category),all_vars(.==sector)) %>% 
    filter(!is.na(WB.income)) %>% 
    filter(Year>=time_start & Year<2018) %>% 
    group_by(Country,Year,WB.income) %>% 
    summarise_at(gas,sum,na.rm=TRUE) %>% 
    ungroup() %>% 
    rename_at(gas,funs(sprintf('value')))
  
  blarg <- left_join(blarg,country_list,by=c("Country"="Country")) %>% 
    filter(cutoff==0) %>% 
    filter(!is.na(WB.income)) %>% 
    filter(Year>=time_start & Year<2018) %>% 
    group_by(WB.income,Year) %>% 
    summarise_at('value',sum,na.rm=TRUE)
  
  blarg <- blarg %>% 
    group_by(WB.income) %>% 
    mutate(rate=(last(value)/first(value))^(1/(last(Year)-time_start+1))-1) %>% 
    mutate(abs_growth=last(value)-first(value)) %>%
    filter(Year==2017) %>% 
    na.omit() %>% 
    mutate(Country=paste(WB.income,"(rest)"))
  
  
  ### join and plot
  rates <- rbind(rates %>% filter(cutoff==1) %>% select(-cutoff),blarg)
  
    rates <- gather(rates %>% mutate(rate=rate*100),var,value,rate:abs_growth)
  
  p2 <- rates %>% 
    filter(var==rate_or_abs) %>% 
    ggplot(.,aes(x = reorder(Country,value),y = value, fill=WB.income)) +
    geom_bar(stat='identity') + 
    ylab(paste(ifelse(rate_or_abs=="rate","Rate of","Absolute")," change in ",gas," emissions (Sector ",sector,") ",time_start,"-2017\n (Countries responsible for 80% of sector emissions are shown)",sep="")) +
    coord_flip() +
    theme_bw() +
    theme(legend.position="none",
          axis.title.y = element_blank()) 

  regions <- left_join(edgar_GHG,country_list,by=c("Country"="Country")) %>%
    filter_at(vars(category),all_vars(.==sector)) %>% 
    filter(!is.na(WB.income)) %>% 
    filter(Country!="World") %>% 
    filter(Year>1979 & Year<2018) %>% 
    filter(cutoff==0) %>% 
    group_by(WB.income,Year) %>% 
    summarise_at(gas,sum,na.rm=TRUE) %>% 
    mutate(Country=paste(WB.income,"(rest)")) %>% 
    ungroup() %>% 
    rename_at(gas,funs(sprintf('value')))
  
  rect <- data.frame(xmin=time_start, xmax=2017, ymin=-Inf, ymax=Inf)
  
  p1 <-regions %>%
    ggplot(.,aes(x=Year,y=value,fill=WB.income)) +
    geom_area() +
    geom_rect(data=rect, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax),
              fill="grey20",
              alpha=0.35,
              inherit.aes = FALSE) +
    theme_bw() +
    theme(legend.position='bottom',
          legend.title=element_blank(),
          axis.title.x = element_blank()) +
    ylab(paste(gas,"Emissions, Gg CO2eq")) +
    ggtitle(title)
  
  
  ############## plot absolute values ############## 
  
  countries <- left_join(edgar_GHG,country_list,by=c("Country"="Country")) %>%
    filter_at(vars(category),all_vars(.==sector)) %>% 
    filter(!is.na(WB.income)) %>% 
    filter(Country!="World") %>% 
    filter(Year>1979 & Year<2018) %>% 
    filter(cutoff==1) %>% 
    group_by(Country,Year,WB.income) %>% 
    summarise_at(gas,sum,na.rm=TRUE) %>% 
    ungroup() %>% 
    rename_at(gas,funs(sprintf('value')))
  
  
  p4 <- ggarrange(p1,p2,ncol=1,nrow=2)
  return(list("plot"=p4,"data"=rates))
}


```

``` {r make_table, include=FALSE}

make_table <- function(data) {
  
  ft <- flextable(data,col_keys=names(data))
  ft <- autofit(ft)
  ft <- fontsize(ft, size = 8,part="all")
  ft <- height_all(ft,0.1)
  ft <- width(ft,j=3,width=5)
  ft <- align(ft,align="left",part="all")
  
  return(ft)
}


```


# 1. All sectors

```{r total_sector_trends,echo=FALSE,warning=FALSE,fig.width=8,fig.height=10,fig.path="../Results/Plots/",dev=c('png','pdf')}

########### create totals ###########
load('../Data/edgar.RData')
totals <- edgar_GHG %>% 
  group_by(ISO,Year,WB.income,Country) %>% 
  summarise(CO2=sum(CO2,rm.na=T),CH4=sum(CH4,rm.na=T),N2O=sum(N2O,rm.na=T),Fgas=sum(Fgas,rm.na=T),GHG=sum(GHG,rm.na=T)) %>% 
  mutate(sector_code="Total",category_final="Total",category_1="Total",category_2="Total",category_3="Total",IPCC_AR5_chapter=0,IPCC_AR5_sector=0) %>% 
  select(ISO,sector_code,category_final,category_1,category_2,category_3,Year,CO2,CH4,N2O,Fgas,GHG,Country,WB.income,IPCC_AR5_chapter,IPCC_AR5_sector)

edgar_GHG <- edgar_GHG %>% 
  bind_rows(totals)

########### prepare plot ###########
p <- edgar_emissions_plot(2010,"sector_code","Total","GHG",'rate',"All sectors, GHG Emissions")
p$plot

########### calc overall trend ###########

trend <- edgar_GHG %>% 
  group_by(Year) %>% 
  summarise(total=sum(GHG))

rm(edgar_GHG)

#write.xlsx(p$data %>% select(-Year),file=paste("../Results/IPCC_plot_data_",Sys.Date(),".xlsx",sep=""),sheetName="Total_GHGs_2010",append=TRUE)


```

# 2. Energy systems (Ch.6)

```{r energy_systems,echo=FALSE,warning=FALSE,fig.width=8,fig.height=10,fig.path="../Results/Plots/",dev=c('png','pdf')}

load('../Data/edgar.RData')

## plot
p <- edgar_emissions_plot(2010,'chapter',6,'GHG','rate','Energy systems, GHG Emissions')
p$plot


categories <- edgar_GHG %>% 
  select(chapter,sector_code,description) %>% 
  distinct(.keep_all = TRUE) %>% 
  filter(chapter==6)
categories$chapter <- as.factor(categories$chapter)

table <- make_table(categories)
table

```

# 2. AFOLU (Ch.7)

```{r AFOLU,echo=FALSE,warning=FALSE,fig.width=8,fig.height=10,fig.path="../Results/Plots/",dev=c('png','pdf')}

load('../Data/edgar.RData')

## plot
p <- edgar_emissions_plot(2010,'chapter',7,'GHG','rate','AFOLU, GHG Emissions')
p$plot


categories <- edgar_GHG %>% 
  select(chapter,sector_code,description) %>% 
  distinct(.keep_all = TRUE) %>% 
  filter(chapter==7)
categories$chapter <- as.factor(categories$chapter)

table <- make_table(categories)
table

#write.xlsx(data %>% select(-Year),file=paste("../Results/IPCC_plot_data_",Sys.Date(),".xlsx",sep=""),sheetName=paste(category,gas,time_start,sep="_"),append=TRUE)


```

# 4. Buildings (Ch.9)

```{r buildings,echo=FALSE,warning=FALSE,fig.width=8,fig.height=10,fig.path="../Results/Plots/",dev=c('png','pdf')}

load('../Data/edgar.RData')

## plot
p <- edgar_emissions_plot(2010,'chapter',9,'GHG','rate','AFOLU, GHG Emissions')
p$plot


categories <- edgar_GHG %>% 
  select(chapter,sector_code,description) %>% 
  distinct(.keep_all = TRUE) %>% 
  filter(chapter==9)
categories$chapter <- as.factor(categories$chapter)

table <- make_table(categories)
table

```

# 5. Transport (Ch.10)

```{r transport,echo=FALSE,warning=FALSE,fig.width=8,fig.height=10,fig.path="../Results/Plots/",dev=c('png','pdf')}

load('../Data/edgar.RData')

## plot
p <- edgar_emissions_plot(2010,'chapter',10,'GHG','rate','AFOLU, GHG Emissions')
p$plot


categories <- edgar_GHG %>% 
  select(chapter,sector_code,description) %>% 
  distinct(.keep_all = TRUE) %>% 
  filter(chapter==10)
categories$chapter <- as.factor(categories$chapter)

table <- make_table(categories)
table

```

# 6. Industry (Ch.11)

```{r industry,echo=FALSE,warning=FALSE,fig.width=8,fig.height=10,fig.path="../Results/Plots/",dev=c('png','pdf')}

load('../Data/edgar.RData')

## plot
p <- edgar_emissions_plot(2010,'chapter',11,'GHG','rate','AFOLU, GHG Emissions')
p$plot


categories <- edgar_GHG %>% 
  select(chapter,sector_code,description) %>% 
  distinct(.keep_all = TRUE) %>% 
  filter(chapter==11)
categories$chapter <- as.factor(categories$chapter)

table <- make_table(categories)
table

```
